Untitled

### 2022-12-13 draft1_Rcode.Rmd
Working on Question 4

@@

    #sampling dataset
    sample.id <- sample(data$id, size = length(y), replace = TRUE)
    #extracting weight and y value corresponding to id sampled
    sample.obs <- NULL
    sample.weight <- NULL
    sample.group <- NULL

    for (i in 1:length(y)){
      sample.obs[i] <- data$Obs[data$id == sample.id[i]]
      sample.weight[i] <- data$weight[data$id == sample.id[i]]
      sample.group[i] <- data$group[data$id == sample.id[i]]
    }
    #creating dataset of sample
    sample.dataset <- data.frame("group"=sample.group,"Obs"=sample.obs,"weight"=sample.weight)

@@

>>>>

```{r}
## Question 4
q4.data <- read.csv("q4MixLinReg.csv",header=T)

#Part a
linRegMixEstep <- function(piCur,aCur,bCur,sCur,x,Y){
  #matrix to store E-Step values
  #row i (p(Z_{i} = 1|Y_{i}, θ^{cur}),..., p(Z_{i} = K|Y_{i}, θ^{cur})) row i has X_{i} ie X fixed
  #column j (p(Z_{i} = j|Y_{1}, θ^{cur}), ...,(Z_{i} = k|Y_{k}, θ^{cur})) column j has Z_{i}=j  ie value of Z_{i} fixed
  prob <- matrix(data=NA,nrow = length(x),ncol = length(piCur))
  for (i in 1:length(x)){
    for (j in 1:length(piCur)){
      prob[i,j] <- piCur[j]*dnorm(Y[i],mean=aCur[j]+bCur[j]*x[i],sd=sCur[j])/sum(piCur*dnorm(Y[i],mean=aCur+bCur*x[i],sd=sCur))
    }
  }
  return(prob)
}

#test <-linRegMixEstep(piCur=c(0.4,0.3,0.3),aCur=c(0.1,-0.1,0.2),bCur=c(1,-1,1),sCur=c(1,0.5,1.1),x=q4.data$x,Y=q4.data$Y)

#Part b
#i
calcNewCoefs <- function(W,x,Y){
  #Calculating a^{new) and b^{new}
  new.coef <- matrix(data=NA,ncol = 2,nrow=dim(W)[2])
  #Creating design matrix
  X <- matrix(c(rep(1,length(x)),x),ncol=2)
  for (k in 1:dim(W)[2]){
    new.coef[k,] <- solve(t(X)%*%diag(W[,k])%*%X)%*%t(X)%*%diag(W[,k])%*%Y
  }
  return(t(new.coef))
}
#new.coef <- calcNewCoefs(test,x=q4.data$x,Y=q4.data$Y)

#ii
calcNewSd <- function(W,x,Y,a,b){
  new.sd <- NULL
  #calculating new sd
  for (k in 1:dim(W)[2]){
    new.sd[k] <- sqrt(sum(W[,k]*(Y-a[k]-b[k]*x)^2)/sum(W[,k]))
  }
  return(new.sd)
}
#new.sd <- calcNewSd(test,x,Y,new.coef[1,],new.coef[2,])

#iii
linRegMixMstep <- function(W,x,Y){
  #calculating new coefficents for a and b
  new.coef <- calcNewCoefs(W,x,Y)

  #calculating new sd
  new.sd <- calcNewSd(W,x,Y,new.coef[1,],new.coef[2,])

  #Calculating new pi
  new.pi <- NULL
  for (k in 1:dim(W)[2]){
    new.pi[k] <- sum(W[,k])/dim(W)[1]
  }
 return(list(new.pi,new.coef[1,],new.coef[2,],new.sd))
}
#new.theta <- linRegMixMstep(test,x,Y)

#c
###############This is wrong!!!!! Need to change (covergence isn't working, plan to use final values from final function in d to see what is happening)
linRegMixCalcLogLik <- function(x,Y,piCur,piNew,aCur,aNew,bCur,bNew,sCur,sNew){
  #calculating conditional probabilities for current and new values of theta
  W.cur <- linRegMixEstep(piCur,aCur,bCur,sCur,x,Y)
  W.new <- linRegMixEstep(piNew,aNew,bNew,sNew,x,Y)
  #Creating object to store values of log-likelihood for each observation i of Y 
  Q.cur <- NULL
  Q.new <- NULL
  for (i in 1:length(Y)){
    #calculating log likelihood for each value of Y
    Q.cur[i] <- sum(W.cur[i,]*log(piCur))+sum(W.cur[i,]*log(dnorm(Y[i],mean = aCur+bCur*x[i],sd=sCur)))
    Q.new[i] <- sum(W.new[i,]*log(piNew))+sum(W.new[i,]*log(dnorm(Y[i],mean = aNew+bNew*x[i],sd=sNew)))
  }
  #Summing over all observations as is in the formula
  Q.cur <- sum(Q.cur)
  Q.new <- sum(Q.new)
  return(c(Q.cur,Q.new))
}

#linRegMixCalcLogLik(x,Y,piCur,new.pi,aCur,new.coef[1,],bCur,new.coef[2,],sCur,new.sd)

#d
linRegMixEM <- function(piInit,aInit,bInit,sInit,x,Y,convergeEps){
  #Initializing the algorithm
  piCur <- piInit
  aCur <- aInit
  bCur <- bInit
  sCur <- sInit

  conv.critera=0
  i=0
  while(i==0){
    #E-step
    W <- linRegMixEstep(piCur,aCur,bCur,sCur,x,Y)
    #Calculating new values for parameters
    theta.new <- linRegMixMstep(W,x,Y)
    #Calculating log-likelihood for current and new values of theta
    theta.loglik <- linRegMixCalcLogLik(x,Y,piCur,theta.new[[1]],aCur,theta.new[[2]],bCur,theta.new[[3]],sCur,theta.new[[4]])
    if(abs(theta.loglik[2]-theta.loglik[1])<convergeEps){
      conv.critera=1
    }

  }
  return(theta.new )
}

test <- linRegMixEM(c(0.4,0.3,0.3),c(0.1,-0.1,0.2),c(1,-1,1),c(1,0.5,1.1),q4.data$x,q4.data$Y,1e-06)

```{r}
linRegMixEM <- function(piInit,aInit,bInit,sInit,x,Y,convergeEps,test){
  #Initializing the algorithm
  piCur <- piInit
  aCur <- aInit
  bCur <- bInit
  sCur <- sInit

  conv.critera=0

  while(i<test){
    #E-step
    W <- linRegMixEstep(piCur,aCur,bCur,sCur,x,Y)
    #Calculating new values for parameters
    theta.new <- linRegMixMstep(W,x,Y)
    #Calculating log-likelihood for current and new values of theta
    theta.loglik <- linRegMixCalcLogLik(x,Y,piCur,theta.new[[1]],aCur,theta.new[[2]],bCur,theta.new[[3]],sCur,theta.new[[4]])

    i=i+1

  }
  return(theta.new )
}

test.1010 <- linRegMixEM(c(0.4,0.3,0.3),c(0.1,-0.1,0.2),c(1,-1,1),c(1,0.5,1.1),q4.data$x,q4.data$Y,1e-06,100)
```

@@

    #sampling dataset
    sample.id <- sample(data$id, size = length(y), replace = TRUE)

    sample.obs <- NULL
    sample.weight <- NULL
    sample.group <- NULL
    #extracting weight, group and y value corresponding to id sampled
    for (i in 1:length(y)){
      sample.obs[i] <- data$Obs[data$id == sample.id[i]]
      sample.weight[i] <- data$weight[data$id == sample.id[i]]
      sample.group[i] <- data$group[data$id == sample.id[i]]
    }
    #creating dataset of sample
    sample.dataset <- data.frame("group"=sample.group,"Obs"=sample.obs,"weight"=sample.weight)

@@

  return(bootstr.se)
}
#Sorry about using so many loops :P
